### 2019-07-11 submission/manuscript.tex
Additional edits before sharing with Pat

Antibiotics are a major risk factor for \emph{Clostridium difficile}
infections (CDIs), but it is less clear what other factors contribute to
CDI incidence and recurrence. Interestingly, other medications besides
antibiotics have been shown to alter the microbiota. Proton pump
inhibitors (PPIs) are associated with alterations in the human
intestinal microbiota through observational and interventional studies
and PPI use has also been associated with CDI incidence and recurrence
in hospital and epidemiological studies. We evaluated the ability of a
high dose of the PPI omeprazole to alter the murine intestinal
microbiota and disrupt colonization resistance to \emph{C. difficile}.
We found daily PPI treatment had minimal impact on the murine intestinal
microbiota and was not enough to promote \emph{C. difficile}
colonization after challenge. Further studies are needed to determine
whether other factors such as the composition of the starting bacterial
community, comorbidities, and use of additional medications contribute
to the association between PPIs and CDIs that is seen in humans.

@@

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium difficile}
infections (CDIs) (1). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between proton pump inhibitor (PPI) use and incidence or
recurrence of CDIs (2--5). PPIs are one of the most widely used drugs in
the U.S. and are available over the counter (6). PPIs are approved to
treat a variety of conditions including gastroesophageal reflux disease,
ulcers, or as prophylaxis for nonsteroidal anti-inflammatory drug
(NSAID)-associated gastrointestinal bleeding (3). Unfortunately, there
have been a number of adverse events associated with their use including
small intestinal bacterial overgrowth, decreased calcium absorption, and
increased risk of enteric infections (3). There have also been a number
of large cohort studies as well as interventional clinical trials that
have demonstrated specific alterations in the intestinal microbiome were
associated with PPI use (4, 7). PPI-associated microbiota changes have
been attributed to the ability of PPIs to increase stomach acid pH which
may promote the survival of oral and pathogenic bacteria (4, 7). Human
fecal microbiota changes with PPI use include increases in
Enterococcaceae, Lactobacillaceae, Micrococcaceae, Staphylococcaceae and
Streptococcaceae and decreases in Ruminococcaceae (7). An \emph{in vitro
screen} demonstrated that 3 PPIs inhibit the growth of 7-8 strains of
Lachnospiraceae and Ruminococcaceae, which represented
\textasciitilde{}16\% of the 40 intestinal bacteria strains that were
screened (8), suggesting direct action on bacteria might also play a
role.\\
However, most of the studies suggesting a link between PPIs and \emph{C.
difficile} were retrospective and did not evaluate the microbiome (2, 3,
5). Thus, it is unclear whether the gastrointestinal microbiome changes
associated with PPI use play a role in the association between PPIs and
CDI incidence or recurrence. Additionally, epidemiological studies have
a limited capacity to address potential confounders and comorbidities in
patients that were on PPIs and developed CDIs or recurrent CDIs (2, 5).
Here, we evaluated the impact of a daily high dose PPI treatment on the
murine microbiome and susceptibility to \emph{C. difficile} colonization
in relation to clindamycin, an antibiotic that perturbs the microbiome
enough to allow \emph{C. difficile} to colonize but is mild enough that
\emph{C. difficile} is cleared within 10 days (9).

\textbf{Murine fecal microbiomes were minimally affected by PPI
treatment } To test if PPI treatment alters the microbiome and promotes
susceptibility to CDIs, we gavaged mice daily with a high dose of
omeprazole for 7 days before \emph{C. difficile} challenge (Figure 1A).
PPI-treated mice were compared to a group of mice that received the
antibiotic, clindamycin, and a group that received both PPI and
clindamycin treatment. A principle coordinates analysis (PCoA) of the
Bray-Curtis distances over the initial 7 days of treatment revealed the
bacterial communities of PPI-treated mice remained relatively unchanged
(Figure 1B). Enterococcaceae, Lactobacillaceae, Micrococcaceae,
Staphylococcaceae, Streptococcaceae, and Ruminococcaceae are all familes
that have previously been impacted by PPIs in human studies, but we
observed no significant fluctuations in the stool of our PPI-treated
mice throughout the course of the 16-day experiment (Figure 1C, Figure
S1). In contrast, the microbiomes from the 2 groups of mice that
received clindamycin started to shift away from the PPI-treated mice and
the samples from previous days one day after treatment (Figure 1D).
These results demonstrated that high dose PPI treatment alone had no
significant impact on the murine fecal bacterial community.

\textbf{PPI-treatment did not promote susceptibility to \emph{C.
difficile infection} in mice} Next, we examined whether high dose PPI
treatment altered susceptibility to \emph{C. difficile} infection in
mice. After 7 days of PPI treatment or 1 day after clindamycin
treatment, mice were challenged with \emph{C. difficile} 630 spores.
While all 4 of the clindamycin-treated mice were colonized with \emph{C.
difficile}, all of the PPI-treated mice were resistant to \emph{C.
difficile} colonization (Figure 2A). Interestingly, only 1 cage of the
PPI and clindamycin-treated mice were colonized, while the other cage
was not (Figure 2A). PCoAs of the bacterial communities from the 3
groups of mice after \emph{C. difficile} challenge, showed that the
greatest shifts in bacterial communities occurred in the
clindamycin-treated mice (Figure 2B). Within 5 days, all of the mice had
cleared \emph{C. difficile}, suggesting there was no difference in rate
of clearance for the mice that were initially colonized with \emph{C.
difficile} despite continuing to receive PPI treatment throughout the
course of the experiment. Our results suggest that PPI treatment alone
had minor effects on microbial community resistance to \emph{C.
difficile} colonization in mice. Instead most of the differences between
our 3 treatment groups appear to be driven by clindamycin administration
and included decreased \emph{Alistipes}, \emph{Barnesiella},
Porhyromonadaceae, Ruminococcaceae, taxa previously found to be altered
in clindamycin-treated mice that were challenged with \emph{C.
difficile} (1). These findings demonstrated that high dose PPI treatment
did not promote susceptibility to \emph{C. difficile} colonization.

@@

PPIs have more of an effect on the small intestinal microbiota compared
to the fecal microbiota (10). We did not find significant changes for
the taxa observed to be significantly impacted by PPI use in human
studies. However, 3 of the families that typically increase were absent
or at low abundance in the mice from the beginning and may be one
contributing factor to why we did not see any changes. Additionally,
some of the significance of PPI associations in human interventional
trials appears to be driven by a handful of specific taxa with overall
differences on a PCoA plot difficult to distinguish (11). Also, most of
the studies to date compareed different individuals (PPI users versus
non-users) or the same individuals before and after treatment (4, 7),
unfortunately these limited microbiota snapshots may be skewed by
day-to-day microbiota variations (13) that would be revealed with more
frequent longitudinal sampling.\\

@@

additional antibiotic treatment, none of these studies looked at the
interplay with the intestinal microbiome (14--16). Additionally, there
were some key methodical differences between the previous studies and
our own. Previous work that administered 0.5 mg/kg of the PPI
lansoprazole daily for 2 weeks to mice and then challenged with \emph{C.
difficile} demonstrated that PPI treatment alone resulted in detectable
\emph{C. difficile} in the stool 1 week after challenge, but also showed
there was detectable \emph{C. difficile} in mice not treated with
antibiotics (14, 15). The presence of \emph{C. difficile} in mice that
were not treated with either antibiotics or PPIs could be attributed to
the higher dose of vegetative cells (10\textsuperscript{8} CFU) used to
challenge the mice and may partly explain the observation of \emph{C.
difficile} in PPI-treated mice (14, 15). We have previously shown mice
from our colony that were not given antibiotics were resistant to
\emph{C. difficile} 630 when challenged with 10\textsuperscript{3}
spores (17). Another group also used a higher dose of
10\textsuperscript{6} spores or 10\textsuperscript{7} vegetative cells
to challenge antibiotic-treated mice or mice treated with both the PPI
esomeprazole (40mg/kg dose) for 2 days and antibiotics and demonstrated
the antibiotic/PPI-treated mice developed more severe CDIs (16).\\

@@

\emph{C. difficile} colonization. Our 16S rRNA sequencing suggested that

@@

C57BL/6 mouse colony. If these genera are important contributors to the
associations between PPIs and CDIs in humans, this may be part of the
reason we did not see an effect in our CDI mouse model. There could also
be differences at the species level for taxa that are altered with PPIs
in humans compared to mice, which may also contribute to the
association. Other murine starting communities or the addition of human
PPI treatment-associated strains may be needed to determine whether PPIs
can alter taxa and subsequently promote CDIs.\\

@@

treatment can increase CDI severity in mice (18) and PPIs are sometimes

@@

\textbf{16S rRNA gene sequencing} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The V4 hypervariable region of the 16S rRNA gene was
amplified with Accuprime Pfx DNA polymerase (Thermo Fisher Scientific)
using previously described custom barcoded primers (19). The 16S rRNA
amplicon library was sequenced with the MiSeq (Illumina). Amplicons were
cleaned up and normalized with the SequalPrep Normalization Plate Kit
(ThermoFisher Scientific) and pooled amplicons were quantified with the
KAPA library quantification kit (KAPA Biosystems).

\textbf{16S rRNA gene sequence analysis}Mothur (v1.40.5) was used for
all sequence processing steps (20) using the previously published
protocol (19). R (v.3.5.1) was used to generate figures and perform
statistical analysis.

@@

\hypertarget{ref-Singh2018}{}
6. \textbf{Singh A}, \textbf{Cresci GA}, \textbf{Kirby DF}. 2018. Proton
pump inhibitors: Risks and rewards and emerging consequences to the gut
microbiome. Nutrition in Clinical Practice \textbf{33}:614--624.
doi:\href{https://doi.org/10.1002/ncp.10181}{10.1002/ncp.10181}.

\hypertarget{ref-Imhann2017}{}
7. \textbf{Imhann F}, \textbf{Vila AV}, \textbf{Bonder MJ},
\textbf{Manosalva AGL}, \textbf{Koonen DP}, \textbf{Fu J},
\textbf{Wijmenga C}, \textbf{Zhernakova A}, \textbf{Weersma RK}. 2017.
The influence of proton pump inhibitors and other commonly used
medication on the gut microbiota. Gut Microbes \textbf{8}:351--358.
doi:\href{https://doi.org/10.1080/19490976.2017.1284732}{10.1080/19490976.2017.1284732}.

\hypertarget{ref-maier2018extensive}{}
8. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{others}. 2018. Extensive impact of
non-antibiotic drugs on human gut bacteria. Nature \textbf{555}:623.

\hypertarget{ref-Jenior2018}{}
9. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. Clostridium difficile alters the structure
and metabolism of distinct cecal microbiomes during initial infection to
promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\hypertarget{ref-Yasutomi2018}{}
10. \textbf{Yasutomi E}, \textbf{Hoshi N}, \textbf{Adachi S},
\textbf{Otsuka T}, \textbf{Kong L}, \textbf{Ku Y}, \textbf{Yamairi H},
\textbf{Inoue J}, \textbf{Ishida T}, \textbf{Watanabe D}, \textbf{Ooi
M}, \textbf{Yoshida M}, \textbf{Tsukimi T}, \textbf{Fukuda S},
\textbf{Azuma T}. 2018. Proton pump inhibitors increase the
susceptibility of mice to oral infection with enteropathogenic bacteria.
Digestive Diseases and Sciences \textbf{63}:881--889.
doi:\href{https://doi.org/10.1007/s10620-017-4905-3}{10.1007/s10620-017-4905-3}.

\hypertarget{ref-Freedberg2015}{}
11. \textbf{Freedberg DE}, \textbf{Toussaint NC}, \textbf{Chen SP},
\textbf{Ratner AJ}, \textbf{Whittier S}, \textbf{Wang TC}, \textbf{Wang
HH}, \textbf{Abrams JA}. 2015. Proton pump inhibitors alter specific
taxa in the human gastrointestinal microbiome: A crossover trial.
Gastroenterology \textbf{149}:883--885.e9.
doi:\href{https://doi.org/10.1053/j.gastro.2015.06.043}{10.1053/j.gastro.2015.06.043}.

\hypertarget{ref-Imhann2015}{}
12. \textbf{Imhann F}, \textbf{Bonder MJ}, \textbf{Vila AV}, \textbf{Fu
J}, \textbf{Mujagic Z}, \textbf{Vork L}, \textbf{Tigchelaar EF},
\textbf{Jankipersadsing SA}, \textbf{Cenit MC}, \textbf{Harmsen HJM},
\textbf{Dijkstra G}, \textbf{Franke L}, \textbf{Xavier RJ},
\textbf{Jonkers D}, \textbf{Wijmenga C}, \textbf{Weersma RK},
\textbf{Zhernakova A}. 2015. Proton pump inhibitors affect the gut
microbiome. Gut \textbf{65}:740--748.
doi:\href{https://doi.org/10.1136/gutjnl-2015-310376}{10.1136/gutjnl-2015-310376}.

\hypertarget{ref-LlornsRico2019}{}
13. \textbf{Lloréns-Rico V}, \textbf{Raes J}. 2019. Tracking humans and
microbes. Nature \textbf{569}:632--633.
doi:\href{https://doi.org/10.1038/d41586-019-01591-y}{10.1038/d41586-019-01591-y}.

\hypertarget{ref-Kaur2007}{}
14. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2007. Comparative role of antibiotic
and proton pump inhibitor in ExperimentalClostridium difficileInfection
in mice. Microbiology and Immunology \textbf{51}:1209--1214.
doi:\href{https://doi.org/10.1111/j.1348-0421.2007.tb04016.x}{10.1111/j.1348-0421.2007.tb04016.x}.

\hypertarget{ref-kaur2011effect}{}
15. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2011. Effect of lactobacillus
acidophilus \& epidermal growth factor on experimentally induced
clostridium difficile infection. The Indian journal of medical research
\textbf{133}:434.

\hypertarget{ref-hung2015proton}{}
16. \textbf{Hung Y-P}, \textbf{Ko W-C}, \textbf{Chou P-H}, \textbf{Chen
Y-H}, \textbf{Lin H-J}, \textbf{Liu Y-H}, \textbf{Tsai H-W}, \textbf{Lee
J-C}, \textbf{Tsai P-J}. 2015. Proton-pump inhibitor exposure aggravates
clostridium difficile--associated colitis: Evidence from a mouse model.
The Journal of infectious diseases \textbf{212}:654--663.

\hypertarget{ref-Jenior2017}{}
17. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. Clostridium difficile colonizes alternative
nutrient niches during infection across distinct murine gut microbiomes.
mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\hypertarget{ref-Maseda2019}{}
18. \textbf{Maseda D}, \textbf{Zackular JP}, \textbf{Trindade B},
\textbf{Kirk L}, \textbf{Roxas JL}, \textbf{Rogers LM},
\textbf{Washington MK}, \textbf{Du L}, \textbf{Koyama T},
\textbf{Viswanathan VK}, \textbf{Vedantam G}, \textbf{Schloss PD},
\textbf{Crofford LJ}, \textbf{Skaar EP}, \textbf{Aronoff DM}. 2019.
Nonsteroidal anti-inflammatory drugs alter the microbiota and exacerbate
clostridium difficile colitis while dysregulating the inflammatory
response. mBio \textbf{10}.
doi:\href{https://doi.org/10.1128/mbio.02282-18}{10.1128/mbio.02282-18}.

\hypertarget{ref-Kozich2013}{}
19. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\hypertarget{ref-Schloss2009}{}
20. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

@@

\includegraphics{results/figures/figure_2.pdf} \textbf{Figure 2. PPI
treatment alone does not promote CDIs in mice} A. C. difficile CFUs/g
stool measured each day post C. difficile challenge for clindamycin,
clindamycin/PPI, and PPI-treated mice. Lines represent the mean CFU/g
for each treatment group while points represent CFU/g for individual
mice within each group. B. PCoA of stool samples collected after
antibiotic treatments. C. Genera significantly associated with treatment
groups for samples across all sequenced timepoints. D. Families
significantly associated with treatment groups for samples across all
sequenced timepoints. C,D data were analyzed by Kruskal-Wallis test with
a Benjamini-Hochberg correction for multiple comparisons.

>>>>

Antibiotics are a major risk factor for \emph{Clostridium difficile}
infections (CDIs), but other factors may also contribute to CDI
incidence and recurrence. Interestingly, other medications besides
antibiotics can impact the microbiota. Proton pump inhibitors (PPIs) are
associated with alterations in the human intestinal microbiota through
observational and interventional studies and PPI use has also been
associated with CDI incidence and recurrence in epidemiological studies.
We evaluated the ability of a high dose of the PPI omeprazole to alter
the murine intestinal microbiota and disrupt colonization resistance to
\emph{C. difficile}. We found daily PPI treatment had minimal impact on
the murine fecal microbiota and was not enough to promote \emph{C.
difficile} colonization after challenge. Further studies are needed to
determine whether other factors such as the composition of the starting
bacterial community, comorbidities, and use of additional medications
contribute to the association between PPIs and CDIs seen in humans.

@@

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium difficile}
infections (CDIs) (1). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between proton pump inhibitor (PPI) use and incidence or
recurrence of CDIs (2--5). There have also been a number of large cohort
studies as well as interventional clinical trials that have demonstrated
specific alterations in the intestinal microbiome were associated with
PPI use (4, 6). PPI-associated microbiota changes have been attributed
to the ability of PPIs to increase stomach acid pH which may promote the
survival of oral and pathogenic bacteria (4, 6). Human fecal microbiota
changes with PPI use include increases in Enterococcaceae,
Lactobacillaceae, Micrococcaceae, Staphylococcaceae and Streptococcaceae
and decreases in Ruminococcaceae (6).\\
Unfortunately, most of the studies suggesting a link between PPIs and
\emph{C. difficile} were retrospective and did not evaluate the
microbiome (2, 3, 5). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use play a role in the
association between PPIs and CDI incidence or recurrence. Additionally,
epidemiological studies have a limited capacity to address potential
confounders and comorbidities in patients that were on PPIs and
developed CDIs or recurrent CDIs (2, 5). Here, we evaluated the impact
of a daily high dose PPI treatment on the murine microbiome and
susceptibility to \emph{C. difficile} colonization in relation to
clindamycin, an antibiotic that perturbs the microbiome enough to allow
\emph{C. difficile} to colonize but is mild enough that \emph{C.
difficile} is cleared within 10 days (7).

\textbf{Murine fecal microbiomes were minimally affected by PPI
treatment } To test if PPI treatment alters the microbiome and promotes
susceptibility to CDIs, we gavaged mice daily with a high dose of
omeprazole for 7 days before \emph{C. difficile} challenge (Figure 1A).
The fecal bacterial communities of PPI-treated mice were compared to a
group of mice that received the antibiotic, clindamycin, and a group
that received both PPI and clindamycin treatment using 16S rRNA gene
sequencing. A principle coordinates analysis (PCoA) of the Bray-Curtis
distances over the initial 7 days of treatment revealed the bacterial
communities of PPI-treated mice remained relatively unchanged (Figure
1B). Enterococcaceae, Lactobacillaceae, Micrococcaceae,
Staphylococcaceae, Streptococcaceae, and Ruminococcaceae are all familes
that have previously been impacted by PPIs in human studies, but we
observed no significant fluctuations in the stool of our PPI-treated
mice throughout the course of the 16-day experiment (Figure 1C, Figure
S1). In contrast, the microbiomes from the 2 groups of mice that
received clindamycin started to shift away from the PPI-treated mice and
the samples from previous days one day after treatment (Figure 1D).
These results demonstrated that high dose PPI treatment alone had no
significant impact on the murine fecal bacterial community.

\textbf{PPI-treatment did not promote susceptibility to \emph{C.
difficile infection} in mice} Next, we examined whether high dose PPI
treatment altered susceptibility to \emph{C. difficile} infection in
mice. After 7 days of PPI treatment or 1 day after clindamycin
treatment, mice were challenged with \emph{C. difficile} 630 spores.
While all 4 of the clindamycin-treated mice were colonized with \emph{C.
difficile}, all of the PPI-treated mice were resistant to \emph{C.
difficile} colonization (Figure 2A). Interestingly, only 1 cage of the
PPI and clindamycin-treated mice were colonized, while the other cage
was resistant (Figure 2A). PCoAs of the bacterial communities from the 3
groups of mice after \emph{C. difficile} challenge, showed that the
greatest shifts in bacterial communities occurred in the
clindamycin-treated mice (Figure 2B). Within 5 days, all of the mice had
cleared \emph{C. difficile}, suggesting there was no difference in rate
of clearance for the mice that were initially colonized with \emph{C.
difficile} despite continuing to receive PPI treatment throughout the
course of the experiment. Our results suggest that PPI treatment alone
had no effects on bacterial community resistance to \emph{C. difficile}
colonization in mice. Instead most of the differences between our 3
treatment groups appear to be driven by clindamycin administration and
included decreased \emph{Alistipes}, \emph{Barnesiella},
Porhyromonadaceae, Ruminococcaceae, taxa previously found to be altered
in clindamycin-treated mice that were challenged with \emph{C.
difficile} (1). These findings demonstrated that high dose PPI treatment
did not promote susceptibility to \emph{C. difficile} colonization.

@@

PPIs had more of an effect on the small intestinal microbiota compared
to the fecal microbiota (8). The same group demonstrated PPI treatment
increased the stomach pH in mice (8), which may improve survival of
bacteria passing through the stomach. We did not find significant
changes for the taxa observed to be significantly impacted by PPI use in
human studies. However, 3 of the families that typically increase were
absent or at low abundance in the mice from the beginning and may be one
contributing factor to our observation that PPIs had no significant
effects on bacterial taxa. Additionally, some of the significance of PPI
associations in human interventional trials appears to be driven by a
handful of specific taxa with overall differences on a PCoA plot
difficult to distinguish (9). Also, most of the studies to date compared
different individuals (PPI users versus non-users) or the same
individuals before and after treatment (4, 6), unfortunately these
limited microbiota snapshots may be skewed by day-to-day microbiota
variations (11) that would be revealed with more frequent longitudinal
sampling.\\

@@

additional antibiotic treatment(12--14), there were some key
methodological differences between these studies and our own. One group
administered 0.5 mg/kg of the PPI lansoprazole daily for 2 weeks to mice
and then challenged with \emph{C. difficile} demonstrated that PPI
treatment alone resulted in detectable \emph{C. difficile} in the stool
1 week after challenge, but also showed there was detectable \emph{C.
difficile} in mice not treated with antibiotics (12, 13). The presence
of \emph{C. difficile} in mice that were not treated with either
antibiotics or PPIs could be attributed to the higher dose of vegetative
cells (10\textsuperscript{8} CFU) used to challenge the mice and may
partly explain the observation of \emph{C. difficile} in PPI-treated
mice (12, 13). We have previously shown mice from our colony that were
not given antibiotics were resistant to \emph{C. difficile} 630 when
challenged with 10\textsuperscript{3} spores (15). The other group also
used a higher dose of 10\textsuperscript{6} spores or
10\textsuperscript{7} vegetative cells to challenge antibiotic-treated
mice or mice treated with both the PPI esomeprazole (40mg/kg dose) for 2
days and antibiotics and demonstrated the antibiotic/PPI-treated mice
developed more severe CDIs (14).\\

@@

\emph{C. difficile} colonization. 16S rRNA sequencing suggested that

@@

C57BL/6 mouse colony. These genera could be important contributors to
the associations between PPIs and CDIs in humans, and could be a
contributing factor to our observation that PPI treatment had no effect
on \emph{C. difficile} colonization in our CDI mouse model. There could
also be differences at the species level for taxa that are altered with
PPIs in humans compared to mice. Other murine starting communities or
the addition of human PPI treatment-associated strains may be needed to
determine whether PPIs can alter taxa and subsequently promote CDIs.\\

@@

treatment can increase CDI severity in mice (16) and PPIs are sometimes

@@

\textbf{16S rRNA gene sequencing} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The V4 hypervariable region of the 16S rRNA gene was
amplified with Accuprime Pfx DNA polymerase (Thermo Fisher Scientific)
using previously described custom barcoded primers (17). The 16S rRNA
amplicon library was sequenced with the MiSeq (Illumina). Amplicons were
cleaned up and normalized with the SequalPrep Normalization Plate Kit
(ThermoFisher Scientific) and pooled amplicons were quantified with the
KAPA library quantification kit (KAPA Biosystems).

\textbf{16S rRNA gene sequence analysis}Mothur (v1.40.5) was used for
all sequence processing steps (18) using the previously published
protocol (17). R (v.3.5.1) was used to generate figures and perform
statistical analysis.

@@

\hypertarget{ref-Imhann2017}{}
6. \textbf{Imhann F}, \textbf{Vila AV}, \textbf{Bonder MJ},
\textbf{Manosalva AGL}, \textbf{Koonen DP}, \textbf{Fu J},
\textbf{Wijmenga C}, \textbf{Zhernakova A}, \textbf{Weersma RK}. 2017.
The influence of proton pump inhibitors and other commonly used
medication on the gut microbiota. Gut Microbes \textbf{8}:351--358.
doi:\href{https://doi.org/10.1080/19490976.2017.1284732}{10.1080/19490976.2017.1284732}.

\hypertarget{ref-Jenior2018}{}
7. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. Clostridium difficile alters the structure
and metabolism of distinct cecal microbiomes during initial infection to
promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\hypertarget{ref-Yasutomi2018}{}
8. \textbf{Yasutomi E}, \textbf{Hoshi N}, \textbf{Adachi S},
\textbf{Otsuka T}, \textbf{Kong L}, \textbf{Ku Y}, \textbf{Yamairi H},
\textbf{Inoue J}, \textbf{Ishida T}, \textbf{Watanabe D}, \textbf{Ooi
M}, \textbf{Yoshida M}, \textbf{Tsukimi T}, \textbf{Fukuda S},
\textbf{Azuma T}. 2018. Proton pump inhibitors increase the
susceptibility of mice to oral infection with enteropathogenic bacteria.
Digestive Diseases and Sciences \textbf{63}:881--889.
doi:\href{https://doi.org/10.1007/s10620-017-4905-3}{10.1007/s10620-017-4905-3}.

\hypertarget{ref-Freedberg2015}{}
9. \textbf{Freedberg DE}, \textbf{Toussaint NC}, \textbf{Chen SP},
\textbf{Ratner AJ}, \textbf{Whittier S}, \textbf{Wang TC}, \textbf{Wang
HH}, \textbf{Abrams JA}. 2015. Proton pump inhibitors alter specific
taxa in the human gastrointestinal microbiome: A crossover trial.
Gastroenterology \textbf{149}:883--885.e9.
doi:\href{https://doi.org/10.1053/j.gastro.2015.06.043}{10.1053/j.gastro.2015.06.043}.

\hypertarget{ref-Imhann2015}{}
10. \textbf{Imhann F}, \textbf{Bonder MJ}, \textbf{Vila AV}, \textbf{Fu
J}, \textbf{Mujagic Z}, \textbf{Vork L}, \textbf{Tigchelaar EF},
\textbf{Jankipersadsing SA}, \textbf{Cenit MC}, \textbf{Harmsen HJM},
\textbf{Dijkstra G}, \textbf{Franke L}, \textbf{Xavier RJ},
\textbf{Jonkers D}, \textbf{Wijmenga C}, \textbf{Weersma RK},
\textbf{Zhernakova A}. 2015. Proton pump inhibitors affect the gut
microbiome. Gut \textbf{65}:740--748.
doi:\href{https://doi.org/10.1136/gutjnl-2015-310376}{10.1136/gutjnl-2015-310376}.

\hypertarget{ref-LlornsRico2019}{}
11. \textbf{Lloréns-Rico V}, \textbf{Raes J}. 2019. Tracking humans and
microbes. Nature \textbf{569}:632--633.
doi:\href{https://doi.org/10.1038/d41586-019-01591-y}{10.1038/d41586-019-01591-y}.

\hypertarget{ref-Kaur2007}{}
12. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2007. Comparative role of antibiotic
and proton pump inhibitor in ExperimentalClostridium difficileInfection
in mice. Microbiology and Immunology \textbf{51}:1209--1214.
doi:\href{https://doi.org/10.1111/j.1348-0421.2007.tb04016.x}{10.1111/j.1348-0421.2007.tb04016.x}.

\hypertarget{ref-kaur2011effect}{}
13. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2011. Effect of lactobacillus
acidophilus \& epidermal growth factor on experimentally induced
clostridium difficile infection. The Indian journal of medical research
\textbf{133}:434.

\hypertarget{ref-hung2015proton}{}
14. \textbf{Hung Y-P}, \textbf{Ko W-C}, \textbf{Chou P-H}, \textbf{Chen
Y-H}, \textbf{Lin H-J}, \textbf{Liu Y-H}, \textbf{Tsai H-W}, \textbf{Lee
J-C}, \textbf{Tsai P-J}. 2015. Proton-pump inhibitor exposure aggravates
clostridium difficile--associated colitis: Evidence from a mouse model.
The Journal of infectious diseases \textbf{212}:654--663.

\hypertarget{ref-Jenior2017}{}
15. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. Clostridium difficile colonizes alternative
nutrient niches during infection across distinct murine gut microbiomes.
mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\hypertarget{ref-Maseda2019}{}
16. \textbf{Maseda D}, \textbf{Zackular JP}, \textbf{Trindade B},
\textbf{Kirk L}, \textbf{Roxas JL}, \textbf{Rogers LM},
\textbf{Washington MK}, \textbf{Du L}, \textbf{Koyama T},
\textbf{Viswanathan VK}, \textbf{Vedantam G}, \textbf{Schloss PD},
\textbf{Crofford LJ}, \textbf{Skaar EP}, \textbf{Aronoff DM}. 2019.
Nonsteroidal anti-inflammatory drugs alter the microbiota and exacerbate
clostridium difficile colitis while dysregulating the inflammatory
response. mBio \textbf{10}.
doi:\href{https://doi.org/10.1128/mbio.02282-18}{10.1128/mbio.02282-18}.

\hypertarget{ref-Kozich2013}{}
17. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\hypertarget{ref-Schloss2009}{}
18. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

@@

\includegraphics{figure_2.pdf} \textbf{Figure 2. PPI treatment alone
does not promote CDIs in mice} A. C. difficile CFUs/g stool measured
each day post C. difficile challenge for clindamycin, clindamycin/PPI,
and PPI-treated mice. Lines represent the mean CFU/g for each treatment
group while points represent CFU/g for individual mice within each
group. B. PCoA of stool samples collected after antibiotic treatments.
C. Genera significantly associated with treatment groups for samples
across all sequenced timepoints. D. Families significantly associated
with treatment groups for samples across all sequenced timepoints. C,D
data were analyzed by Kruskal-Wallis test with a Benjamini-Hochberg
correction for multiple comparisons.
